### 2023-01-18 manuscript.Rmd
some refs added

Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes (Schönbrodt et al., X), Gaertner et al. suggest a specific implementation to broaden the range of what constitutes research contributions. Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions. Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3). In the following, we comment on dimensions metrics, lines of code, software engineering, impact, peer-review, usability.

@@

Gaertner et al. suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package. They acknowledge the problem that citations inconsistently point to software packages and/or accompanying software packages. For example, one of the authors maintains the pdc package for R. Google scholar counts X citations to the package and X citations to the accompanying paper, of which XX citations cite both, which permits a simple addition of the citation counts.

@@

In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note that these metrics are difficult to compare across different types of software package. However, we would like to point out that the request for citation metrics with respect to software packages seems inconsistent with the authors condemnation of citation metrics for publications.

Complementing the quantitative aspect of impact, we suggest to introduce a qualitative response. Researchers should indicate at least one (and up to three) empirical research projects that directly benefitted from the software contributions they made. We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.

>>>>

Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes (Schönbrodt et al., X), @gartner2022responsible suggest a specific implementation to broaden the range of what constitutes research contributions. Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions. Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3). In the following, we comment on the dimensions metrics, lines of code, software engineering, impact, peer-review, usability.

@@

@gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package. They acknowledge the problem that citations inconsistently point to software packages and/or accompanying software packages. For example, one of the authors maintains the pdc package for R. Google scholar counts X citations to the package and X citations to the accompanying paper, of which XX citations cite both, which permits a simple addition of the citation counts.

@@

In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note that these metrics are difficult to compare across different types of software packages. For example, it is probably more commonly accepted to cite data analysis packages (particularly if they are commercial packages) than supporting packages (such as papaja [@R-papaja] used to render this article). A further complication arises through the fact that functionality of successful, modular scientific software is imported, re-used, and re-exported in other software packages. While we highly encourage this from both a software engineering perspective and for scientific progress, it poses further difficulties to citation-based metrics. For example, the pdc package [@pdc-paper] offers functions to cluster time-series based on relative complexity. The TSclust package [@TSclust] is a wrapper package that unites several divergence measures for time series clustering conveniently in a single package, particularly, the pdc clustering approach. This potentially leads to researchers only citing the TSclust package but not the original algorithm (e.g., as in the work by @staroverova).

Last, we would like to point out that the request for citation metrics with respect to software packages seems inconsistent with the authors condemnation of citation metrics for publications.

Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what the others proposed as indicator "merit / impact statement". In our view, researchers should be requested to indicate at least one (and up to three) empirical research projects that directly benefitted from the software contributions they made. We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.
